Prediction of 90-day mortality after Total Hip Arthroplasty: a simplified and externally validated model based on observational registry data from Sweden, England and Wales

### 2020-03-27 reports/tabs.Rmd
version with table 2 combining SHAR and NJR. Updated Shiny app.

<!-- FIGURES -->

```{r flowchart, fig.cap = "Flowchart depicting inclusion criteria and number of patients. Data from the Swedish Hip Arthroplasty Register was used for model derivation and internal validation (left). Data from the National Joint Registry in England and Wales was used for external validation (right)."}
knitr::include_graphics("../graphs/flowchart.png")
```

@@

```{r rocs, fig.cap = "Receiver Operation Characteristics (ROC) curves. The area under the ROC curve (AUC) for the reduced model was no different compared to the main model (left panel). The reduced model works almost as good for predictions in the National Joint Registry for England and Wales (right panel)."}
knitr::include_graphics("../graphs/roc.png")
```

```{r aucci, fig.cap = "Area Under the Receiver Operation Characteristics Curve (AUC) as a measure of predictive discriminative ability with 95 % bootstrap confidence intervals. The reduced model was no different compared to the main model, neither internally, nor externally (overlapping confidence intervals). AUC above 0.7 was considered good. This was observed for the derived models but not for any other model included for comparison."}

knitr::include_graphics("../graphs/auc_ci.png")
```

```{r sep, fig.cap = "Most patients survived more than 90 days after THA, wherefore the blue bars totally dominate the histogram (upper panel; note the scales). A normalized density plot however reveals that patients who died within 90 days were, on average, estimated to have a higher probability to do so (lower panel). There were very few observations with covariate patterns resulting in death probabilities higher than 5% (green line). Estimated probabilities above this limit are therefore subject to extrapolation."}
knitr::include_graphics("../graphs/separation.png")
```

```{r calibration, fig.cap = "This figure illustrates calibrartion between observed proportions and predicted probabilities with 95 % confidence intervals. Deviations above the diagonal line means that the model underpredicts the probability of death within 90 days. Such deviation was statistically significant for the external validation for predictions below 1.5%. The higher the predicted probability gets, the more they over-estimate the observed proportions of deaths."}
knitr::include_graphics("../graphs/calibration.png")
```

<!-- TABLES -->

```{r tabcategorization}
knitr::kable(
  tab_categorization, 
  caption = "Categorization of co-morbdities based on diagnostic groups according to Charlson and Elixhauser. (CNS = central nervous system.)")

@@

```{r tab1}
knitr::kable(
  table1, 
  caption = "Characteristics of the study population. Co-morbidities recorded for at least one patient who survived 90 days, and one who did not, were modeled as potential predictors. (CNS = central nervous system.)")
```

@@

<!-- Same info in figure so do not use this table -->
```{r auc, warning = FALSE, include = FALSE}
knitr::kable(
  brlasso_tbl_auc %>% 
  filter(
    `Age as` %in% c(NA, "main effect"),
    !endsWith(Model, "cancer") 
  ) %>% 
  select(Model, AUC) %>% 
  rename(`AUC (95 % CI)` = AUC),
  caption = "Area Under the Curve (AUC) as a measure of discriminative ability for the main and reduced models, a model with age and sex, as well as univariable models with ASA class and the Charlson or Elixhauser co-morbidity indices.")
```

@@

```{r brlreduced}
brlasso_tbl_coefs %>% 
  select(-math) %>%
  mutate(
    term = gsub("= 2", "II", term),
    term = gsub("= 3", "III", term)
  ) %>% 
knitr::kable(
  caption = 'Estimated coefficients (beta) and odds ratios (OR) with 95% confidence intervals for the reduced model.'
)
```

>>>>

@@

```{r tabcategorization}
tab_categorization %>%
  ungroup() %>% 
  mutate(
    `Comorbidities by groups` = 
      gsub("Heart infarct", "Myocardial infarction", `Comorbidities by groups`)
  ) %>% 
  knitr::kable(
    caption = "Categorization of co-morbidities based on diagnostic groups according to Charlson and Elixhauser. (CNS = central nervous system.)")

@@

```{r tab1}
knitr::kable(
  mutate(table1_combined_tot, what = gsub("Heart infarct", "Myocardial infarction", what)), 
  caption = "Characteristics of the study population in the model derivation cohort (SHAR) and external validation cohort (NJR). Educational levels were classified as low (up to 9 years), middle (10-12 years) and high (at least 12 years).  BMI = Body mass index. ASA = American Society for Anesthesiologists classification. CNS = central nervous system.")
```

@@

@@

```{r brlreduced}
brlasso_tbl_coefs %>% 
  select(-math) %>%
  mutate(
    term = gsub("= 2", "II", term),
    term = gsub("= 3", "III", term)
  ) %>% 
knitr::kable(
  caption = 'Estimated coefficients (beta) and odds ratios (OR) with 95 % confidence intervals for the reduced model.'
)
```
